---
title: "MoPP Examples in the Beta-Bernoulli Model"
author: "Marc Box"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
bibliography: "references.bib"
vignette: >
  %\VignetteIndexEntry{MoPP Examples in the Beta-Bernoulli Model}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# MoPP Examples in the Beta-Bernoulli Model

The beta-Bernoulli model is 

This is an example used in Scott et al 2016. Binary data is simulated and we 
infer the rate parameter of a Bernoulli model, specifying a beta distribution 
pior.

The posterior distribution in Scott et al 2016's example is positively skewed, 
and so non-normal, presenting a particular challenge for the Consensus 
algorithm. We will show that this is not a challenge for the MoPP algorithm, 
which we contend is because we are able to use the intended, and proper, prior 
distribution rather than the "fractionated" version.

First we will look at the specific example of Scott et al 2016. Then we will 
look at a more general example with a different rate parameter.

## The example from Scott et al 2016

Scott et al 2016's data set consists of 1,000 binary observations including a 
single success and 999 failures. These are partitioned into 100 data shards.

This is an interesting test for distributed computation because the success 
outcome will only be observed in one shard. The question is how well are we 
able to reproduce the infereces we would have made if all 1,000 observations 
had been accessible to our posterior sampling routine.

```{r}
n <- 1000
n_success <- 1
x <- c(rep(0, n - n_success), rep(1, n_success))
```

Partition the data into $M = 100$ shards.

```{r}
M <- 100
x.list <- partition(x, M, random = TRUE, balanced = TRUE)
```

### Option: distribute data shards to Spark cluster

This block is not run. It merely demonstrates how the data could be distributed 
using Spark.

```{r eval = FALSE}
# NOT RUN
library(sparklyr)

config <- spark_config()
config$spark.executor.memory <- "1125m"
config$spark.executor.cores <- 1
config$spark.executor.instances <- 2
config$spark.dynamicAllocation.enabled <- FALSE

config$`sparklyr.cores.local` <- 8
config$`sparklyr.shell.driver-memory` <- "8G"
config$spark.dynamicAllocation.enabled <- "false"
config$spark.executor.instances <- M

sc <- spark_connect(
  master = "local",
  config = config
)

invoke(spark_context(sc), "getExecutorMemoryStatus")

x.cluster.parts <- list()
for (p in 1:M) x.cluster.parts[[p]] <- sdf_copy_to(sc, as.data.frame(cbind(p, x.list[[p]])), repartition = 1, overwrite = TRUE)
# This creates a Spark table with M "partitions" (i.e. distributed parts).
x.cluster <- do.call(sdf_bind_rows, x.cluster.parts)

# Check a few things.
spark_apply(x.cluster, function(p) nrow(p), names = "n")
spark_apply(x.cluster, function(p) tail(p, 2))

# Check number of partitions.
sdf_num_partitions(x.cluster)
```

## Sample from the posterior distribution

Here we draw 10,000 samples from the full posterior distribution and 
$\frac{10,000}{M}$ from each of the partial posterior distributions. There are 
three partial posterior distributions we could draw from, defined by their 
different prior distributions. These are:

1. The unmodified beta prior distribution described above.
2. The fractionated version of the prior distribution proposed by Scott et al 
2016.
3. The fractionated version of the prior distribution obtained by taking the 
$M^\textrm{th}$ root of the unnormalised beta prior density.

When the beta prior is $\text{rm{Beta}(1,1)$, the third option is identical to 
the first. So in this example there are only two partial posteriors.

The entropy (in nats) of the beta prior is

```{r}
lbeta(a, b) - (a - 1) * digamma(a) - (b - 1) * digamma(b) + (a + b - 2) * digamma(a + b)
```

Here we sample from the full posterior distribution.

```{r}
H <- 10000
# Beta prior parameters.
a <- 1
b <- 1

a.post <- a + sum(x)
b.post <- b + n - sum(x)
samples.full <- rbeta(H, a.post, b.post)
```

Here we sample from the partial posterior defined with the unmodified beta 
prior.

```{r}
samples.part <- lapply(x.list, FUN = function(xi) {
    a.post = a + sum(xi)
    b.post = b + length(xi) - sum(xi)
    matrix(rbeta(ceiling(H / M), a.post, b.post), ceiling(H / M), 1)
})
```

And here we sample from the partial posterior defined with the fractionated 
prior.

```{r}
samples.part.frac <- lapply(x.list, FUN = function(xi) {
    a.post = a / M + sum(xi)
    b.post = b / M + length(xi) - sum(xi)
    matrix(rbeta(ceiling(H / M), a.post, b.post), ceiling(H / M), 1)
})
```

This is what the true posterior looks like.

```{r}
x.grid <- qbeta(log(seq(1e-6, 1 - 1e-6, by = 1e-6)), a.post, b.post, log.p = TRUE)
posterior_plot_data <- data.frame(
  Lambda = c(x.grid),
  Density = dbeta(x.grid, a.post, b.post),
  Series = rep("Posterior", length(x.grid))
)
ggplot2::ggplot(posterior_plot_data, ggplot2::aes(x = Lambda, y = Density)) +
ggplot2::geom_line() +
ggplot2::scale_x_continuous(name = expression(lambda)) +
ggplot2::theme_bw(base_size = 8) +
ggplot2::theme(panel.grid.minor = ggplot2::element_blank())
```

The posterior mean is

```{r}
a.post / (a.post + b.post)
```

And the posterior variance is

```{r}
a.post * b.post / (a.post + b.post) ^ 2 / (a.post + b.post + 1)
```

Plotting the samples.

```{r}
dens <- density(samples.full)
plot_data <- rbind(posterior_plot_data, data.frame(
  Lambda = dens$x[dens$x >= 0 & dens$x <= 1],
  Density = dens$y[dens$x >= 0 & dens$x <= 1],
  Series = rep("Posterior samples", sum(dens$x >= 0 & dens$x <= 1))
))
ggplot2::ggplot(plot_data, ggplot2::aes(x = Lambda, y = Density, linetype = Series)) +
ggplot2::geom_line() +
ggplot2::scale_x_continuous(name = expression(lambda)) +
ggplot2::scale_linetype_manual(values = c("Posterior" = "solid", "Posterior samples" = "dotted")) +
ggplot2::theme_bw(base_size = 8) +
ggplot2::theme(panel.grid.minor = ggplot2::element_blank())
```


## Pooling algorithms

Here we define the log likelihood function, to be used by MoPP.

```{r}
loglik.fun <- function(theta, x) {
  n <- length(x)
  n_pos <- sum(x)
  n_pos * log(theta) + (n - n_pos) * log(1 - theta)
}
```

MoPP.

```{r}
timer <- proc.time()
mopp.out <- mopp.weights(
  x.list,
  samples.part,
  loglik.fun = loglik.fun,
  type = 2,
  keep.type1 = TRUE
)
print(proc.time() - timer)
```

Consensus Monte Carlo algorithm.

```{r}
timer <- proc.time()
consensus1.out <- consensus.weights(
  samples.part.frac,
  type = 1,
  return.pooled = TRUE
)
print(proc.time() - timer)

timer <- proc.time()
consensus2.out <- consensus.weights(
  samples.part.frac,
  type = 2,
  return.pooled = TRUE
)
print(proc.time() - timer)
```

Density product estimators.

```{r}
timer <- proc.time()
ndpe.out <- dpe.master(
  samples.part.frac,
  type = "ndpe",
  keep.type1 = TRUE
)
print(proc.time() - timer)

timer <- proc.time()
sdpe.out <- dpe.master(
  samples.part.frac,
  type = "sdpe",
  keep.type1 = TRUE
)
print(proc.time() - timer)
```

Plot the MoPP and consensus algorithm approximations of the posterior density.

```{r}
dens <- density(consensus2.out$theta.w.pooled)
plot_data <- rbind(posterior_plot_data, data.frame(
  Lambda = dens$x[dens$x >= 0 & dens$x <= 1],
  Density = dens$y[dens$x >= 0 & dens$x <= 1],
  Series = rep("Consensus (weighted)", sum(dens$x >= 0 & dens$x <= 1))
))

x.grid <- seq(quantile(samples.full, prob = 1e-4), quantile(samples.full, prob = 1 - 1e-4), by = dens$bw)
dens.mopp <- mopp.kde(
  x.grid,
  samples.part,
  mopp.out$wn.type_2,
  bw = dens$bw
)
plot_data <- rbind(plot_data, data.frame(
  Lambda = x.grid,
  Density = dens.mopp,
  Series = rep("MoPP (type 2)", length(x.grid))
))

ggplot2::ggplot(plot_data, ggplot2::aes(x = Lambda, y = Density, linetype = Series)) +
ggplot2::geom_line() +
ggplot2::scale_x_continuous(name = expression(lambda)) +
ggplot2::scale_linetype_manual(values = c("Posterior" = "solid", "Consensus (weighted)" = "dotted", "MoPP (type 2)" = "dashed")) +
ggplot2::theme_bw(base_size = 8) +
ggplot2::theme(panel.grid.minor = ggplot2::element_blank())
```

Here is a quantile-quantile plot for the consensus and MoPP algorithms.

```{r}
# True posterior quantiles.
p.grid <- c(
  exp(seq(log(1e-6), log(0.5), by = 1e-4)),
  1 - exp(seq(log(1e-6), log(0.5), by = 1e-4))
)
q.post <- qbeta(p.grid, a.post, b.post)

# Consensus estimates.
qq_data <- data.frame(
  Posterior_Quantile = q.post,
  Estimated_Quantile = quantile(consensus2.out$theta.w.pooled, probs = p.grid),
  Series = rep("Consensus (weighted)", length(p.grid))
)

# MoPP estimates.
p.mopp <- cumsum(exp(unlist(mopp.out$wn.type_2)[order(unlist(samples.part))]))
q.mopp <- sort(unlist(samples.part))[findInterval(p.grid, p.mopp, all.inside = TRUE)]
qq_data <- rbind(qq_data, data.frame(
  Posterior_Quantile = q.post,
  Estimated_Quantile = q.mopp,
  Series = rep("MoPP (type 2)", length(p.grid))
))

# Add diagonal?
# qq_data <- rbind(qq_data, data.frame(
#  Posterior_Quantile = q.post,
#  Estimated_Quantile = q.post,
#  Series = rep("Identity", length(p.grid))
#)

ggplot2::ggplot(qq_data, ggplot2::aes(x = Posterior_Quantile, y = Estimated_Quantile, linetype = Series)) +
ggplot2::geom_line() +
ggplot2::scale_x_continuous(name = "Posterior quantile") +
ggplot2::scale_y_continuous(name = "Estimated quantile") +
ggplot2::scale_linetype_manual(values = c("Consensus (weighted)" = "dotted", "MoPP (type 2)" = "dashed")) +
ggplot2::theme_bw(base_size = 8) +
ggplot2::theme(
  panel.grid.minor = ggplot2::element_blank(),
  panel.grid.major = ggplot2::element_blank()
  legend.position= c(1 - 0.8, 0.75),
  legend.title = ggplot2::element_blank()
)
```


### Posterior parameter estimates

Here we compute the pooling algorithm estimates of the posterior mean, median, 
2.5% quantile, 97.5% quantile and standard deviation of each model parameter 
dimension.

We also estimate the effective sample size (ESS). "Roughly speaking, the ESS of 
a quantity of interest captures how many independent draws contain the same 
amount of information as the dependent sample obtained by the MCMC algorithm" 
(Vehtari et al 2019). The maximum possible for the consensus and DPE algorithms 
is $\frac{H^{total}}{M}$; for the other methods, including MoPP, the maximum 
possible is $H$.

```{r}
posterior.estimates <- list()
ess <- sapply(c("Vanilla", "Naive", "MoPP 1", "MoPP 2", "Consensus 1", "Consensus 2", "NDPE", "SDPE"), FUN = function(a) {NA})
```

Full posterior sampling.

```{r}
posterior.estimates[["Vanilla"]] <- c(
  "Mean" = mean(samples.full),
  "Median" = median(samples.full),
  quantile(samples.full, 0.025),
  quantile(samples.full, 0.975),
  "Std. dev." = sd(samples.full)
)
ess["Vanilla"] <- mcmcse::ess(samples.full)
```

Naive pooling of partial posterior samples.

```{r}
posterior.estimates[["Naive"]] <- apply(
  abind::abind(samples.part, along = 1),
  2,
  FUN = function(theta) {
    c(
      "Mean" = mean(theta),
      "Median" = median(theta),
      quantile(theta, 0.025),
      quantile(theta, 0.975),
      "Std. dev." = sd(theta)
    )
  }
)
ess["Naive"] <- mcmcse::ess(unlist(samples.part))
```

MoPP algorithm 1.

```{r}
posterior.estimates[["MoPP 1"]] <- rbind(
  "Mean" = mopp.mean(samples.part, mopp.out$wn.type_1, type = 1, Hvec = mopp.out$Hvec),
  "Median" = mopp.quantile(samples.part, 0.5, mopp.out$wn.type_1, type = 1, Hvec = mopp.out$Hvec)["root",],
  "2.5%" = mopp.quantile(samples.part, 0.025, mopp.out$wn.type_1, type = 1, Hvec = mopp.out$Hvec)["root",],
  "97.5%" = mopp.quantile(samples.part, 0.975, mopp.out$wn.type_1, type = 1, Hvec = mopp.out$Hvec)["root",]
)
posterior.estimates[["MoPP 1"]] <- rbind(
  posterior.estimates[["MoPP 1"]],
  "Std. Dev" = sqrt(mopp.mean(lapply(samples.part, FUN = function(theta) {theta ^ 2}), mopp.out$wn.type_1, type = 1, Hvec = mopp.out$Hvec) - posterior.estimates[["MoPP 1"]]["Mean",] ^ 2)
)

ess["MoPP 1"] <- exp(2 * log(sum(mopp.out$Hvec)) -
  lrowsums(
    unlist(
      mapply(
        FUN = function(w, h) {2 * w + 2 * log(h)},
        mopp.out$wn.type_1,
        mopp.out$Hvec,
        SIMPLIFY = FALSE
      )
    )
  )
)
```

MoPP algorithm 2.

```{r}
posterior.estimates[["MoPP 2"]] <- rbind(
  "Mean" = mopp.mean(samples.part, mopp.out$wn.type_2, type = 2, Hvec = mopp.out$Hvec),
  "Median" = mopp.quantile(samples.part, 0.5, mopp.out$wn.type_2, type = 2, Hvec = mopp.out$Hvec)["root",],
  "2.5%" = mopp.quantile(samples.part, 0.025, mopp.out$wn.type_2, type = 2, Hvec = mopp.out$Hvec)["root",],
  "97.5%" = mopp.quantile(samples.part, 0.975, mopp.out$wn.type_2, type = 2, Hvec = mopp.out$Hvec)["root",]
)
posterior.estimates[["MoPP 2"]] <- rbind(
  posterior.estimates[["MoPP 2"]],
  "Std. Dev" = sqrt(mopp.mean(lapply(samples.part, FUN = function(theta) {theta ^ 2}), mopp.out$wn.type_2, type = 2, Hvec = mopp.out$Hvec) - posterior.estimates[["MoPP 2"]]["Mean",] ^ 2)
)

ess["MoPP 2"] <- exp(-lrowsums(2 * unlist(mopp.out$wn.type_2)))
```

Consensus Monte Carlo.

```{r}
posterior.estimates[["Consensus 1"]] <- apply(
  consensus1.out$theta.w.pooled,
  2,
  FUN = function(theta) {
    c(
      "mean" = mean(theta),
      "median" = median(theta),
      quantile(theta, 0.025),
      quantile(theta, 0.975),
      "std. dev." = sd(theta)
    )
  }
)

ess["Consensus 1"] <- mcmcse::ess(consensus1.out$theta.w.pooled)

posterior.estimates[["Consensus 2"]] <- apply(
  consensus2.out$theta.w.pooled,
  2,
  FUN = function(theta) {
    c(
      "mean" = mean(theta),
      "median" = median(theta),
      quantile(theta, 0.025),
      quantile(theta, 0.975),
      "std. dev." = sd(theta)
    )
  }
)

ess["Consensus 2"] <- mcmcse::ess(consensus2.out$theta.w.pooled)
```

Density product estimators.

```{r}
posterior.estimates[["NDPE"]] <- apply(
  ndpe.out$theta.pooled,
  2,
  FUN = function(theta) {
    c(
      "mean" = mean(theta),
      "median" = median(theta),
      quantile(theta, 0.025),
      quantile(theta, 0.975),
      "std. dev." = sd(theta)
    )
  }
)

ess["NDPE"] <- mcmcse::ess(ndpe.out$theta.pooled)

posterior.estimates[["SDPE"]] <- apply(
  sdpe.out$theta.pooled,
  2,
  FUN = function(theta) {
    c(
      "mean" = mean(theta),
      "median" = median(theta),
      quantile(theta, 0.025),
      quantile(theta, 0.975),
      "std. dev." = sd(theta)
    )
  }
)

ess["SDPE"] <- mcmcse::ess(sdpe.out$theta.pooled)
```



## Example with $\lambda = 0.5$

I contend that the fractionated prior used in the previous example gives rise 
to acceptable results with the consensus algorithm only because it happens to 
be similar to the posterior distribution (density concentrated near zero).

What if we used the same prior distribution in a similar example but where the 
true success rate is 0.5?

```{r}
n <- 1000
n_success <- n / 2
x <- c(rep(0, n - n_success), rep(1, n_success))
```

Partition the data into $M = 100$ shards.

```{r}
M <- 100
x.list <- partition(x, M, random = TRUE, balanced = TRUE)
```

### Option: distribute data shards to Spark cluster

```{r eval = FALSE}
# NOT RUN
x.cluster.parts <- list()
for (p in 1:M) x.cluster.parts[[p]] <- sdf_copy_to(sc, as.data.frame(cbind(p, x.list[[p]])), repartition = 1, overwrite = TRUE)
# This creates a Spark table with M "partitions" (i.e. distributed parts).
x.cluster <- do.call(sdf_bind_rows, x.cluster.parts)
```

## Sample from the posterior distribution

Here we sample from the full posterior distribution.

```{r}
H <- 10000
# Beta prior parameters.
a <- 1
b <- 1

a.post <- a + sum(x)
b.post <- b + n - sum(x)
samples.full <- rbeta(H, a.post, b.post)
```

Here we sample from the partial posterior defined with the unmodified beta 
prior.

```{r}
samples.part <- lapply(x.list, FUN = function(xi) {
    a.post = a + sum(xi)
    b.post = b + length(xi) - sum(xi)
    matrix(rbeta(ceiling(H / M), a.post, b.post), ceiling(H / M), 1)
})
```

And here we sample from the partial posterior defined with the fractionated 
prior.

```{r}
samples.part.frac <- lapply(x.list, FUN = function(xi) {
    a.post = a / M + sum(xi)
    b.post = b / M + length(xi) - sum(xi)
    matrix(rbeta(ceiling(H / M), a.post, b.post), ceiling(H / M), 1)
})
```

This is what the true posterior looks like.

```{r}
x.grid <- qbeta(log(seq(1e-6, 1 - 1e-6, by = 1e-6)), a.post, b.post, log.p = TRUE)
posterior_plot_data <- data.frame(
  Lambda = c(x.grid),
  Density = dbeta(x.grid, a.post, b.post),
  Series = rep("Posterior", length(x.grid))
)
ggplot2::ggplot(posterior_plot_data, ggplot2::aes(x = Lambda, y = Density)) +
ggplot2::geom_line() +
ggplot2::scale_x_continuous(name = expression(lambda)) +
ggplot2::theme_bw(base_size = 8) +
ggplot2::theme(panel.grid.minor = ggplot2::element_blank())
```

The posterior mean is

```{r}
a.post / (a.post + b.post)
```

And the posterior variance is

```{r}
a.post * b.post / (a.post + b.post) ^ 2 / (a.post + b.post + 1)
```

Plotting the samples.

```{r}
dens <- density(samples.full)
plot_data <- rbind(posterior_plot_data, data.frame(
  Lambda = dens$x[dens$x >= 0 & dens$x <= 1],
  Density = dens$y[dens$x >= 0 & dens$x <= 1],
  Series = rep("Posterior samples", sum(dens$x >= 0 & dens$x <= 1))
))
ggplot2::ggplot(plot_data, ggplot2::aes(x = Lambda, y = Density, linetype = Series)) +
ggplot2::geom_line() +
ggplot2::scale_x_continuous(name = expression(lambda)) +
ggplot2::scale_linetype_manual(values = c("Posterior" = "solid", "Posterior samples" = "dotted")) +
ggplot2::theme_bw(base_size = 8) +
ggplot2::theme(panel.grid.minor = ggplot2::element_blank())
```


## Pooling algorithms

MoPP.

```{r}
timer <- proc.time()
mopp.out <- mopp.weights(
  x.list,
  samples.part,
  loglik.fun = loglik.fun,
  type = 2,
  keep.type1 = TRUE
)
print(proc.time() - timer)
```

Consensus Monte Carlo algorithm.

```{r}
timer <- proc.time()
consensus1.out <- consensus.weights(
  samples.part.frac,
  type = 1,
  return.pooled = TRUE
)
print(proc.time() - timer)

timer <- proc.time()
consensus2.out <- consensus.weights(
  samples.part.frac,
  type = 2,
  return.pooled = TRUE
)
print(proc.time() - timer)
```

Density product estimators.

```{r}
timer <- proc.time()
ndpe.out <- dpe.master(
  samples.part.frac,
  type = "ndpe"
)
print(proc.time() - timer)

timer <- proc.time()
sdpe.out <- dpe.master(
  samples.part.frac,
  type = "sdpe"
)
print(proc.time() - timer)
```

Plot the MoPP and consensus algorithm approximations of the posterior density.

```{r}
dens <- density(consensus2.out$theta.w.pooled)
plot_data <- rbind(posterior_plot_data, data.frame(
  Lambda = dens$x[dens$x >= 0 & dens$x <= 1],
  Density = dens$y[dens$x >= 0 & dens$x <= 1],
  Series = rep("Consensus (weighted)", sum(dens$x >= 0 & dens$x <= 1))
))

x.grid <- seq(quantile(samples.full, prob = 1e-4), quantile(samples.full, prob = 1 - 1e-4), by = dens$bw)
dens.mopp <- mopp.kde(
  x.grid,
  samples.part,
  mopp.out$wn.type_2,
  bw = dens$bw
)
plot_data <- rbind(plot_data, data.frame(
  Lambda = x.grid,
  Density = dens.mopp,
  Series = rep("MoPP (type 2)", length(x.grid))
))

ggplot2::ggplot(plot_data, ggplot2::aes(x = Lambda, y = Density, linetype = Series)) +
ggplot2::geom_line() +
ggplot2::scale_x_continuous(name = expression(lambda)) +
ggplot2::scale_linetype_manual(values = c("Posterior" = "solid", "Consensus (weighted)" = "dotted", "MoPP (type 2)" = "dashed")) +
ggplot2::theme_bw(base_size = 8) +
ggplot2::theme(panel.grid.minor = ggplot2::element_blank())
```

Here is a quantile-quantile plot for the consensus and MoPP algorithms.

```{r}
# True posterior quantiles.
p.grid <- c(
  exp(seq(log(1e-6), log(0.5), by = 1e-4)),
  1 - exp(seq(log(1e-6), log(0.5), by = 1e-4))
)
q.post <- qbeta(p.grid, a.post, b.post)

# Consensus estimates.
qq_data <- data.frame(
  Posterior_Quantile = q.post,
  Estimated_Quantile = quantile(consensus2.out$theta.w.pooled, probs = p.grid),
  Series = rep("Consensus (weighted)", length(p.grid))
)

# MoPP estimates.
p.mopp <- cumsum(exp(unlist(mopp.out$wn.type_2)[order(unlist(samples.part))]))
q.mopp <- sort(unlist(samples.part))[findInterval(p.grid, p,mopp, all.inside = TRUE)]
qq_data <- rbind(qq_data, data.frame(
  Posterior_Quantile = q.post,
  Estimated_Quantile = q.mopp,
  Series = rep("MoPP (type 2)", length(p.grid))
))

# Add diagonal?
# qq_data <- rbind(qq_data, data.frame(
#  Posterior_Quantile = q.post,
#  Estimated_Quantile = q.post,
#  Series = rep("Identity", length(p.grid))
#)

ggplot2::ggplot(qq_data, ggplot2::aes(x = Posterior_Quantile, y = Estimated_Quantile, linetype = Series)) +
ggplot2::geom_line() +
ggplot2::scale_x_continuous(name = "Posterior quantile") +
ggplot2::scale_y_continuous(name = "Estimated quantile") +
ggplot2::scale_linetype_manual(values = c("Consensus (weighted)" = "dotted", "MoPP (type 2)" = "dashed")) +
ggplot2::theme_bw(base_size = 8) +
ggplot2::theme(
  panel.grid.minor = ggplot2::element_blank(),
  panel.grid.major = ggplot2::element_blank()
  legend.position= c(1 - 0.8, 0.75),
  legend.title = ggplot2::element_blank()
)
```


### Posterior parameter estimates

Here we compute the pooling algorithm estimates of the posterior mean, median, 
2.5% quantile, 97.5% quantile and standard deviation of each model parameter 
dimension.

```{r}
posterior.estimates <- list()
ess <- sapply(c("Vanilla", "Naive", "MoPP 1", "MoPP 2", "Consensus 1", "Consensus 2", "NDPE", "SDPE"), FUN = function(a) {NA})
```

Full posterior sampling.

```{r}
posterior.estimates[["Vanilla"]] <- apply(
  samples.full,
  2,
  FUN = function(theta) {
    c(
      "Mean" = mean(theta),
      "Median" = median(theta),
      quantile(theta, 0.025),
      quantile(theta, 0.975),
      "Std. dev." = sd(theta)
    )
  }
)

ess["Vanilla"] <- mcmcse::ess(samples.full)
```

Naive pooling of partial posterior samples.

```{r}
posterior.estimates[["Naive"]] <- apply(
  abind::abind(samples.part, along = 1),
  2,
  FUN = function(theta) {
    c(
      "Mean" = mean(theta),
      "Median" = median(theta),
      quantile(theta, 0.025),
      quantile(theta, 0.975),
      "Std. dev." = sd(theta)
    )
  }
)

ess["Naive"] <- mcmcse::ess(unlist(samples.part))
```

MoPP algorithm 1.

```{r}
posterior.estimates[["MoPP 1"]] <- rbind(
  "Mean" = mopp.mean(samples.part, mopp.out$wn.type_1, type = 1, Hvec = mopp.out$Hvec),
  "Median" = mopp.quantile(samples.part, 0.5, mopp.out$wn.type_1, type = 1, Hvec = mopp.out$Hvec)["root",],
  "2.5%" = mopp.quantile(samples.part, 0.025, mopp.out$wn.type_1, type = 1, Hvec = mopp.out$Hvec)["root",],
  "97.5%" = mopp.quantile(samples.part, 0.975, mopp.out$wn.type_1, type = 1, Hvec = mopp.out$Hvec)["root",]
)
posterior.estimates[["MoPP 1"]] <- rbind(
  posterior.estimates[["MoPP 1"]],
  "Std. Dev" = sqrt(mopp.mean(lapply(samples.part, FUN = function(theta) {theta ^ 2}), mopp.out$wn.type_1, type = 1, Hvec = mopp.out$Hvec) - posterior.estimates[["MoPP 1"]]["Mean",] ^ 2)
)

ess["MoPP 1"] <- exp(2 * log(sum(mopp.out$Hvec)) -
  lrowsums(
    unlist(
      mapply(
        FUN = function(w, h) {2 * w + 2 * log(h)},
        mopp.out$wn.type_1,
        mopp.out$Hvec,
        SIMPLIFY = FALSE
      )
    )
  )
)
```

MoPP algorithm 2.

```{r}
posterior.estimates[["MoPP 2"]] <- rbind(
  "Mean" = mopp.mean(samples.part, mopp.out$wn.type_2, type = 2, Hvec = mopp.out$Hvec),
  "Median" = mopp.quantile(samples.part, 0.5, mopp.out$wn.type_2, type = 2, Hvec = mopp.out$Hvec)["root",],
  "2.5%" = mopp.quantile(samples.part, 0.025, mopp.out$wn.type_2, type = 2, Hvec = mopp.out$Hvec)["root",],
  "97.5%" = mopp.quantile(samples.part, 0.975, mopp.out$wn.type_2, type = 2, Hvec = mopp.out$Hvec)["root",]
)
posterior.estimates[["MoPP 2"]] <- rbind(
  posterior.estimates[["MoPP 2"]],
  "Std. Dev" = sqrt(mopp.mean(lapply(samples.part, FUN = function(theta) {theta ^ 2}), mopp.out$wn.type_2, type = 2, Hvec = mopp.out$Hvec) - posterior.estimates[["MoPP 2"]]["Mean",] ^ 2)
)

ess["MoPP 2"] <- exp(-lrowsums(2 * unlist(mopp.out$wn.type_2)))
```

Consensus Monte Carlo.

```{r}
posterior.estimates[["Consensus 1"]] <- apply(
  consensus1.out$theta.w.pooled,
  2,
  FUN = function(theta) {
    c(
      "mean" = mean(theta),
      "median" = median(theta),
      quantile(theta, 0.025),
      quantile(theta, 0.975),
      "std. dev." = sd(theta)
    )
  }
)

ess["Consensus 1"] <- mcmcse::ess(consensus1.out$theta.w.pooled)

posterior.estimates[["Consensus 2"]] <- apply(
  consensus2.out$theta.w.pooled,
  2,
  FUN = function(theta) {
    c(
      "mean" = mean(theta),
      "median" = median(theta),
      quantile(theta, 0.025),
      quantile(theta, 0.975),
      "std. dev." = sd(theta)
    )
  }
)

ess["Consensus 2"] <- mcmcse::ess(consensus2.out$theta.w.pooled)
```

Density product estimators.

```{r}
posterior.estimates[["NDPE"]] <- apply(
  ndpe.out$theta.pooled,
  2,
  FUN = function(theta) {
    c(
      "mean" = mean(theta),
      "median" = median(theta),
      quantile(theta, 0.025),
      quantile(theta, 0.975),
      "std. dev." = sd(theta)
    )
  }
)

ess["NDPE"] <- mcmcse::ess(ndpe.out$theta.pooled)

posterior.estimates[["SDPE"]] <- apply(
  sdpe.out$theta.pooled,
  2,
  FUN = function(theta) {
    c(
      "mean" = mean(theta),
      "median" = median(theta),
      quantile(theta, 0.025),
      quantile(theta, 0.975),
      "std. dev." = sd(theta)
    )
  }
)

ess["SDPE"] <- mcmcse::ess(sdpe.out$theta.pooled)
```





Close cluster, if used.

```{r eval = FALSE}
# NOT RUN
spark_disconnect(sc)
```

